---
title: "The Kingdom of God Is Within You"
author: "Jamie Hargreaves"
output: rmarkdown::github_document
---

<style>
  body{
    font-size: 14px;
  }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction and Setting Up

In this project, we'll use tools from the tidyverse to do some simple exploratory analysis (followed by some sentiment analysis), of "The Kingdom of God is Within You", by Leo Tolstoy. We'll be using [Project Gutenberg](http://www.gutenberg.org/) as our data source.

#### Loading libraries

Let's load our libraries and download the data. Note that whilst we won't explicitly load the library, we'll need to make sure that the cowplot package is installed so we can use of the ```cowplot::plot_grid()``` function later on.

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(tidytext)
library(magrittr)
library(gutenbergr)
library(syuzhet)
library(ggraph)
```

#### Importing our data

We can see the list of books available from Project Gutenberg using the ```gutenberg_metadata``` function from the gutenbergr package:

```{r}
gutenberg_metadata
```

We'll use dplyr and stringr to look for any books similar to our title, since there might be multiple versions available from Project Gutenberg, and download using the book ID

```{r message = FALSE}
gutenberg_metadata %>%
  filter(str_detect(str_to_lower(title), "kingdom of god")) %>%
  select(title, gutenberg_id)

book <- gutenberg_download(43302)
book %<>%
  select(-one_of("gutenberg_id"))

book %>%
  head(12)
```

#### Pre-processing

Before we do anything else, we need to clean the text up a little. We can see already that we have several rows containing only ```""``` and we also look to have a preface and other text preceeding the first chapter of the book. We'll add chapter and line numbers and filter out anything before the first chapter:

```{r}
book %<>%
  filter(text != "") %>%
  mutate(
    chapter = cumsum(
      str_detect(text, regex("chapter [\\divxlc]", ignore_case = TRUE))
    )
  ) %>%
  filter(chapter > 0) %>%
  mutate(line = row_number()) %>%
  select(chapter, line, text)

book %>%
  head()
```

We should also look at the last lines in the book to make sure we haven't included any appendices or similar sections that might distort our analysis:

```{r}
book %>%
  arrange(desc(line))
```

The book actually ends on line 11,200 so let's get rid of any lines after that:

```{r}
book %<>%
  filter(line <= 11200)
```

Now we'll tokenise the text:

```{r}
book_tokenised <- book %>%
  unnest_tokens(word, text)
```

Project Gutenberg uses ```_``` to denote words which are italicised, so we're likely to have some of these words in our text:

```{r}
book_tokenised %>%
  filter(str_detect(word, "_"))
```

We've returned some instances of underscored words, so let's remove them:

```{r}
book_tokenised %<>%
  mutate(word = str_replace_all(word, "_", ""))
```

Now that our book is tokenised, we need to remove any stop-words (like "I", "and", "to" etc.):

```{r message = FALSE}
data(stop_words)

book_tokenised %<>%
  anti_join(stop_words)
```


## Single Word Analysis

The data is in a tidy format, so we can start to analyse it. Let's look at the top ten words:

```{r message = FALSE, fig.align = "center", fig.retina = 2}
book_tokenised %>%
  count(word, sort = TRUE) %>%
  top_n(10) %>%
  ggplot(aes(x = reorder(word, n), y = n)) + 
  geom_col() + 
  coord_flip() + 
  xlab(NULL) + 
  ylab("Frequency")
```

We can also see the top words per chapter. For the sake of readability we'll restrict ourselves to the top 3:

```{r message = FALSE, fig.align = "center", fig.width = 13, fig.retina = 2}
book_tokenised %>%
  group_by(chapter) %>%
  count(word, sort = TRUE) %>%
  arrange(chapter) %>%
  top_n(3) %>%
  ggplot(aes(x = word, y = n, fill = as.factor(chapter))) +
  geom_col(show.legend = FALSE) + 
  coord_flip() +
  facet_grid(~chapter, scales = "free_x") + 
  xlab(NULL) + 
  ylab("Frequency") + 
  theme(axis.text.x = element_text(angle = 45))
```


## Analysing $n$-grams

We can get further insight by analysing $n$-grams, i.e. sequences of words. We can tokenise our book into bigrams using tidytext:

```{r}
book_bigrams <- book %>%
  select(chapter, text) %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
  mutate(bigram = str_replace_all(bigram, "_", ""))

book_bigrams %>%
  head()
```

Let's split the bigrams into two separate columns:

```{r}
book_bigrams_separated <- book_bigrams %>%
  separate(bigram, c("word1", "word2"), sep = " ")

book_bigrams_separated %>%
  head()
```

And now let's remove stop-words and take a look at the most frequent bigrams:

```{r message = FALSE, fig.align = "center", fig.retina = 2}
book_bigrams_filtered <- book_bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)

book_bigrams_filtered %>%
  unite(bigram, word1, word2, sep = " ") %>%
  count(bigram, sort = TRUE) %>%
  top_n(10) %>%
  ggplot(aes(x = reorder(bigram, n), y = n)) +
  geom_col() +
  coord_flip() + 
  xlab(NULL) + 
  ylab("Frequency")
```

To visualise how these terms are connected, we'll use ggraph to produce a network graph of common bigrams:

```{r fig.align = "center", fig.width = 8, fig.retina = 2}
set.seed(100)

book_bigrams_filtered %>%
  count(word1, word2, sort = TRUE) %>%
  filter(n > 10) %>%
  igraph::graph_from_data_frame() %>%
  ggraph(layout = "kk") + 
  geom_edge_link(
    aes(width = n), 
    alpha = 0.3, 
    colour = "orange", 
    show.legend = FALSE
  ) + 
  geom_node_point(colour = "orange2", size = 5) + 
  geom_node_text(aes(label = name), vjust = 1, hjust = -0.25, size = 3) + 
  coord_cartesian(clip = "off") + 
  theme_graph()
```

The thicker the connection between two words, the more frequently those words appeared together. We can see that a word like "christian" forms a central node reflecting its prominence throughout the book, and we can see the prominence of topics like "military service" and "christ's teaching". Whilst this is fairly obvious for a book whose themes we already know, these networks can be extremely powerful when trying to piece together themes from things like survey results or tweets.

### Sentiment Analysis

Next, let's look at performing some sentiment analysis. We'll use the ```AFINN``` lexicon to assign a sentiment score to each word :

```{r message = FALSE}
afinn <- get_sentiments("afinn")

book_sentiments <- book_tokenised %>%
  inner_join(afinn)
```

We want to calculate the net sentiment over a sensible period; we could look at net sentiment per chapter, but it's unlikely to tell us anything genuinely meaningful since a lot of the nuance would likely be lost. Instead, we'll look at the net sentiment per "paragraph" which I'm going to define to be every 50 words (remembering that we've taken out a lot of stop-words, so our "paragraphs" should probably be shorter than a physical paragraph):

```{r fig.retina = 2, fig.align = "center"}
book_sentiments %>%
  mutate(paragraph = row_number()%/%50) %>%
  group_by(paragraph) %>%
  summarise(netscore = sum(score)) %>%
  ggplot(aes(x = paragraph, y = netscore)) + 
  geom_col() + 
  xlab("Paragraph") + 
  ylab("Net Sentiment Score")
```

As an aside, we can create a smoothed approximation of the net sentiment score to get a clearer idea of the general sentiment trend as the book goes on:

```{r fig.retina = 2, fig.width = 13}
# net sentiment score per paragraph
sentiment_per_paragraph <- book_sentiments %>%
  mutate(paragraph = row_number()%/%50) %>%
  group_by(paragraph) %>%
  summarise(netscore = sum(score)) 

# original plot
net_sentiment_raw <- sentiment_per_paragraph %>%
  ggplot(aes(x = paragraph, y = netscore)) + 
  geom_col() + 
  xlab("Paragraph") + 
  ylab(NULL)

# transformed plot
smooth_sentiment <- get_dct_transform(sentiment_per_paragraph[[2]], low_pass_size = 15, 
                                      scale_vals = TRUE, 
                                      x_reverse_len = nrow(sentiment_per_paragraph))

book_theme <- data.frame(paragraph = 1:length(smooth_sentiment), score = smooth_sentiment) %>%
  as_tibble()

net_sentiment_trans <- book_theme %>%
  ggplot(aes(x = paragraph, y = score)) +
  geom_col(width = 0.3, fill = "orange") + 
  geom_line(colour = "orange2", size = 1.1) + 
  xlab("Paragraph") + 
  ylab(NULL)

cowplot::plot_grid(net_sentiment_raw, net_sentiment_trans, labels = NULL)
```

Whilst the above is insightful, it doesn't take into account context. When we remove stop-words, we also remove negations, meaning phrases like "no good" and "not happy" are truncated to "good" and "happy", and (misleadingly) area assigned a positive sentiment score. Let's look at words preceeded by negations such as "no", "not" or "never":

```{r}
negations <- c("no", "not", "never")
stop_words_filtered <- stop_words %>%
  filter(!word %in% negations)

bigram_negations <- book_bigrams_separated %>%
  filter(!word1 %in% stop_words_filtered$word) %>%
  filter(!word2 %in% stop_words$word) %>%
  filter(word1 == "no" | word1 == "not" | word1 == "never") 

bigram_negations %>%
  count(word1, word2, sort = TRUE) %>%
  head(10)
```

We can guage the extent to which these words have "mislead" us by looking at the overall contribution of each word to the net sentiment, i.e. the word's sentiment score multiplied by the number of times it occurs:

```{r}
negation_contribution <- bigram_negations %>%
  count(word2, sort = TRUE) %>%
  inner_join(afinn, by = c(word2 = "word")) %>%
  mutate(contribution = n * score) %>%
  arrange(desc(abs(contribution)))
  
negation_contribution
```

Rather than a bar chart, let's use a "lollipop" plot to visualise the contribution:

```{r fig.align = "center", fig.retina = 2}
negation_contribution %>%
  head(20) %>%
  ggplot(aes(x = reorder(word2, contribution), y = contribution)) + 
  geom_segment(aes(x = reorder(word2, contribution), 
                      xend = 
                      reorder(word2,contribution), 
                      y = 0, 
                      yend = contribution), 
                      colour = "gray") +
  geom_point(aes(colour = contribution > 0), show.legend = FALSE, size = 3.5) +
  coord_flip() + 
  xlab("Words Preceeded by Negation") + 
  ylab("Contribution (n*score)") +
  theme(
    plot.background = element_blank(),
    panel.background = element_rect(fill = "gray94"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(colour = "lightgray"),
    panel.grid.minor.x = element_line(colour = "lightgray", linetype = "dashed")
  )
```

We can see that negations of "evil", kill" and "free" are responsible for a large amount of the false sentiment scoring.