---
title: "The Kingdom of God Is Within You"
author: "Jamie Hargreaves"
date: "23 March, 2019"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up
#### Loading libraries

We'll start by loading the relevant libraries and importing the data. We'll be using tidyverse for dplyr, ggplot2 and stringr, tidytext as the basis for our text mining, magrittr for its pipes (mainly ```%<>%```, since ```%>%``` is already loaded with tidyverse), gutenbergr for our book and syuzhet for low pass filtering. Note that whilst we won't load the library, we'll need to also make sure that the cowplot package is installed, as we'll be making use of the ```cowplot::plot_grid()``` function later on.

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(tidytext)
library(magrittr)
library(gutenbergr)
library(syuzhet)
```

#### Importing our data

We'll be using [Project Gutenberg](http://www.gutenberg.org/) as the source of our text data. We can see the list of books available from Project Gutenberg using the ```gutenberg_metadata``` function from the gutenbergr package:

```{r}
gutenberg_metadata
```

We can use dplyr and stringr to look for any books similar to our title, since there might be multiple versions available from Project Gutenberg:

```{r}
gutenberg_metadata %>%
  filter(str_detect(str_to_lower(title), "kingdom of god")) %>%
  select(title, gutenberg_id)
```

We want the second book in the list, so we'll pass its ID to the ```gutenberg_download``` function:

```{r, message = FALSE}
book <- gutenberg_download(43302)
book %<>%
  select(-one_of("gutenberg_id"))

book %>%
  head(12)
```

### Pre-processing

Before we do anything else, we need to clean the text up a little. We can see already that we have several rows containing only ```""``` and we also look to have a preface and other text preceeding the first chapter of the book. We'll add chapter and line numbers to our tibble and filter out anything before the first chapter. For the line number we'll use the ```row_number()``` function from dplyr, and to work out which chapter each line is in we'll use a cumulative sum to count the occurrences of a regular expression ```"chapter [\\divxlc]"```, i.e. "chapter" followed by a number or Roman numeral (any combination of "I", "V", "X", "L" or "C"):

```{r}
book %<>%
  filter(text != "") %>%
  mutate(
    chapter = cumsum(
      str_detect(text, regex("chapter [\\divxlc]", ignore_case = TRUE))
    )
  ) %>%
  filter(chapter > 0) %>%
  mutate(line = row_number()) %>%
  select(chapter, line, text)

book %>%
  head()
```

We should also look at the last lines in the book to make sure we haven't included any appendices or similar sections that might distort our analysis:

```{r}
book %>%
  arrange(desc(line))
```

We can see that the book actually ends on line 11,200, so we'll remove any lines after that:

```{r}
book %<>%
  filter(line <= 11200)
```

We can now use tidytext's ```unnest_tokens()``` function to tokenise our text, i.e. convert it to a one-word-per-row tibble:

```{r}
book_tokenised <- book %>%
  unnest_tokens(word, text)
```

Project Gutenberg uses ```_``` to denote words which are italicised, so we're likely to have some of these words in our tokenised tibble:

```{r}
book_tokenised %>%
  filter(str_detect(word, "_"))
```

We've returned some instances of underscored words, so we'll remove them using a regular expression in ```str_extract()``` from the stringr package:

```{r}
book_tokenised %<>%
  mutate(word = str_extract(word, "[\\da-z]+"))
```

Now that our book is tokenised, we need to remove any stop words (like "I", "and", "to" etc.). Luckily, the tidytext package includes a stop-words data set, and we can use ```anti_join()``` from dplyr along with our tokenised tibble to remove them:

```{r message = FALSE}
data(stop_words)

book_tokenised %<>%
  anti_join(stop_words)
```


### Word Analysis

Now that our data is in a tidy format, we can easily start to analyse it using tools from the tidyverse. Let's start by looking at the overall top ten words using dplyr's ```top_n()``` function and ggplot2:

```{r message = FALSE, fig.align = "center", fig.retina = 2}
book_tokenised %>%
  count(word, sort = TRUE) %>%
  top_n(10) %>%
  ggplot(aes(x = reorder(word, n), y = n)) + 
  geom_col() + 
  coord_flip() + 
  xlab(NULL) + 
  ylab("Frequency")
```

We can also see the top words per chapter; for the sake of readability we'll restrict ourselves to the top 3 words per chapter:

```{r message = FALSE, fig.width = 13, fig.retina = 2}
book_tokenised %>%
  group_by(chapter) %>%
  count(word, sort = TRUE) %>%
  arrange(chapter) %>%
  top_n(3) %>%
  ggplot(aes(x = word, y = n, fill = as.factor(chapter))) +
  geom_col(show.legend = FALSE) + 
  coord_flip() +
  facet_grid(~chapter, scales = "free_x") + 
  xlab(NULL) + 
  ylab("Frequency") + 
  theme(axis.text.x = element_text(angle = 45))
```

### Sentiment Analysis

Next, let's look at performing an analysis of the sentiments of The Kingdom of God Is Within You. We'll use the ```AFINN``` lexicon to assign a sentiment score to each word in our tokenised data frame. The tidytext package includes a useful ```get_sentiments()``` function to easily grab the lexicon, and we can use the ```inner_join()``` function to score each word:

```{r message = FALSE}
afinn <- get_sentiments("afinn")

book_sentiments <- book_tokenised %>%
  inner_join(afinn)

book_sentiments %>%
  head()
```

We want to calculate the net sentiment over a sensible period; we could look at net sentiment per chapter, but it's unlikely to tell us anything truly meaningful since a lot of the nuance would likely be lost. Instead, we'll look at the net sentiment per "paragraph" which I'm going to define to be every 50 words (remembering that we've taken out a lot of stop-words, so our "paragraphs" should probably be somewhat shorter than a typical, physical paragraph).

```{r fig.retina = 2, fig.align = "center"}
book_sentiments %>%
  mutate(paragraph = row_number()%/%50) %>%
  group_by(paragraph) %>%
  summarise(netscore = sum(score)) %>%
  ggplot(aes(x = paragraph, y = netscore)) + 
  geom_col() + 
  xlab("Paragraph") + 
  ylab("Net Sentiment Score")
```

As an aside, we can create a smoothed approximation of the net sentiment score to get a clearer idea of the general sentiment trend as the book goes on (we'll use the syuzhet package for this):

```{r fig.retina = 2, fig.width = 13}
# net sentiment score per paragraph
sentiment_per_paragraph <- book_sentiments %>%
  mutate(paragraph = row_number()%/%50) %>%
  group_by(paragraph) %>%
  summarise(netscore = sum(score)) 

# original plot
net_sentiment_raw <- sentiment_per_paragraph %>%
  ggplot(aes(x = paragraph, y = netscore)) + 
  geom_col() + 
  xlab("Paragraph") + 
  ylab(NULL)

# transformed plot
smooth_sentiment <- get_dct_transform(sentiment_per_paragraph[[2]], low_pass_size = 15, 
                                      scale_vals = TRUE)

book_theme <- data.frame(paragraph = 1:length(smooth_sentiment), score = smooth_sentiment) %>%
  as_tibble()

net_sentiment_trans <- book_theme %>%
  ggplot(aes(x = paragraph, y = score)) +
  geom_col(width = 0.2, fill = "orange") + 
  geom_line(colour = "orange2", size = 1.1) + 
  xlab("Paragraph") + 
  ylab(NULL)

cowplot::plot_grid(net_sentiment_raw, net_sentiment_trans, labels = NULL)
```